Experience Report:

### 2010-03-15 doc/icfp-2010/main.tex
*** empty log message ***

\usepackage{amsmath}

@@

@@

main idea of the master function

four scenarios:
\begin{itemize}
\item cores
\item same binary
\item same version of ocaml
\item completely different programs
\end{itemize}

@@

@@

>>>>

\usepackage{amsmath,graphicx, color}

\newcommand{\Ocaml}{OCaml}
\newcommand{\unix}{\textsc{Unix}}

@@

master and worker terminology introduced here

a few technical details about marshaling should be explained

@@

The main idea is to start with a generic function, with the
following interface: 
% TODO: rename this function
\begin{ocaml}
  val compute : 
    f:('a -> 'b) -> 
    handle:('a * 'c -> 'b -> ('a * 'c) list) -> 
    ('a * 'c) list -> unit
\end{ocaml}
This function takes two functions, \ocaml{f} and \ocaml{handle}, and a
list of \emph{tasks} of type \ocaml{('a * 'c) list}.
The function \ocaml{f} is applied to the first component for each task.
Whenever it completes, the task and the result are passed to the
function \ocaml{handle}. This may possibly generate a new list of
tasks. The process ends when all computations are done.

Function \ocaml{f} should be pure and is executed in parallel.
Function \ocaml{handle}, on the contrary, can be impure and is never
executed in parallel with other \ocaml{handle}s. Typically,
\ocaml{handle} stores and processes information locally.
The next section will describe the usage of this generic interface to
perform traditional map and fold operations.

Actually, our library provides more than just a single \ocaml{compute}
function as above, but instead provides five different versions
depending on the execution context. The five possible contexts are the
following: 
\begin{enumerate}
\item \textbf{purely sequential execution:}
  this is intended for validating the results;

\item \textbf{several cores on the same machine:} 
  this implementation makes use of \unix\ processes and thus can
  provide a polymorphic \ocaml{compute} function;

\item \textbf{same executable run on master and worker machines:}
  this implementation makes use of the ability to marshal \Ocaml\
  closures and thus can still provide a polymorphic \ocaml{compute} function;

\item \textbf{master and workers are different programs, compiled with
    the same version of \Ocaml:} 
  in this case, we can no longer marshal closures but we can still
  marshal polymorphic values. As a consequence, 
  the \ocaml{compute} function is split into two
  polymorphic functions, to implement the master and workers respectively:
  \begin{ocaml}
val master : 
  handle:('a * 'c -> 'b -> ('a * 'c) list) -> 
  ('a * 'c) list -> unit
val worker : ('a -> 'b) -> unit -> unit
  \end{ocaml}

\item \textbf{master and workers are different programs, not even
    compiled with the same version of \Ocaml:} in this case, we can no
  longer use marshaling, so the
  \ocaml{compute} function is split into two monomorphic functions
  over strings: 
  \begin{ocaml}
val master : 
  handle:(string * 'c -> string -> (string * 'c) list) -> 
  (string * 'c) list -> unit
val worker : (string -> string) -> unit -> unit
  \end{ocaml}
  Note that the second component of each task is still polymorphic (of
  type \ocaml{'c} here).
\end{enumerate}

Our library is organized into three modules: \of{Sequential} for the
pure sequential implementation, \of{Cores} for multiple cores on the
same machine and \of{Network} for a network of machines, respectively.
The \of{Network} module itself is organized into three sub-modules,
corresponding to contexts 3, 4 and 5 above. The different
API functions are summarized in the following table (their signatures
being given above):
\begin{center}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    \ocaml{Sequential} & \ocaml{Cores} &
    \multicolumn{3}{|c|}{\ocaml{Network}} 
    \\\cline{3-5}
    &       & \ocaml{Same} & \ocaml{Poly} & \ocaml{Mono} \\\hline
    \ocaml{compute}  & \ocaml{compute}   & \ocaml{compute}  &
    \ocaml{master}  & \ocaml{master}  \\
    & & & \ocaml{worker}  & \ocaml{worker} \\\hline
  \end{tabular}
\end{center}

@@

In most cases, the easiest way to parallelize execution it to make use
of map and fold operations over lists, where processing of the list
elements are done in parallel.  Using the generic API described in the
previous section to do that is not very convenient.
Hence we provide the most commonly used list operations in our API.

The most obvious operation is the traditional map operation over
lists, that is:
\begin{ocaml}
  val map : ('a -> 'b) -> 'a list -> 'b list
\end{ocaml}
The next natural operation is a combination of map and fold
operations, that is a function like
\begin{ocaml}
 val map_fold :
   map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
   'c -> 'a list -> 'c
\end{ocaml}
which, given two functions, an accumulator \of{a} and a list \of{l}, computes
\begin{equation}\label{eq:map-fold}
  \of{fold} ... (\of{fold} (\of{fold} ~ a ~ (\of{map} ~ x_1)) (\of{map} ~ x_2))
  ... (\of{map} ~ x_n)
\end{equation}
for some permutation $[x_1,x_2,...,x_n]$ of the list \of{l}.

We assume that the \of{map} operations are always performed in parallel.
Regarding \of{fold} operations, we distinguish two cases:
\begin{itemize}
\item either \of{fold} operations are computationally less expensive
  than \of{map} and we perform them locally;
\item or \of{fold} operations are computationally expensive and we
  perform them in parallel.
\end{itemize}
We thus provide two functions \of{map_local_fold} and \of{map_remote_fold}.

In the case of \of{map_remote_fold}, only one \of{fold} operation can
be performed at a time (possibly in parallel with \of{map}
operations), as obvious from~(\ref{eq:map-fold}). However, there are
cases where several \of{fold} operations can be performed in parallel,
as early as intermediate results of \of{fold} operations are available.
This is the case when \of{fold} is an associative operation (which
implies that type \of{'b} and \of{'c} are the same  and that the
accumulator should be a neutral element for \of{fold}). Whenever
\of{fold} is also commutative, we can perform even more \of{fold}
operations in parallel. Thus our API provides two functions
\of{map_fold_a} and \of{map_fold_ac} for these two particular cases.
The five operations of the derived API are summarized
in Figure~\ref{fig:derived}.
\begin{figure}[t]
  \begin{ocaml}
    val map : 
      f:('a -> 'b) -> 'a list -> 'b list 
    val map_local_fold : 
      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
      'c -> 'a list -> 'c 
    val map_remote_fold : 
      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
      'c -> 'a list -> 'c 
    val map_fold_ac : 
      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
      'b -> 'a list -> 'b 
    val map_fold_a : 
      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
      'b -> 'a list -> 'b
  \end{ocaml}
  \caption{Derived API.}
\label{fig:derived}
\end{figure}

These functions are actually derived from the generic API.

@@

% LocalWords:  parallelize
